Stack Safety is a Security Property

### 2021-04-17 tex/main.tex

%\bcp{Does use of unitialized variables lead to stack safety attacks?\apt{sure, via buffer overflow}}
%with the notion usually being
%defined negatively: through examples of stack-based exploits
%\rb{usually though not always; as noted below there are other
%  approaches}\bcp{Yes, we have to be very careful about this}.
%\leo{Examples at a high level}: such as -- does not overwrite return addresses, or ...
The most abstract characterization of stack safety that we are aware of is
recent work by \citet{SkorstengaardSTK}, which defines stack safety as the conjunction of two
properties:
{\em local state encapsulation (LSE)} and {\em well-bracketed control flow (WBCF)}.
Informally, LSE says that the contents of
a caller's stack frame are not read or written while its callees (and their callees, etc.) are
executing, and WBCF says that callees always
return to the instruction following their call (if they return at all).
Skorstengaard et al.{} formalize these properties by defining an idealized
machine in which they hold {\em by construction};
they then show that an enforcement
mechanism (in their case, hardware capabilities) guarantees stack safety, by
proving that an ordinary machine with enforcement simulates the idealized
machine.
%an enforcement mechanism (in their case, hardware
%capabilities)
%can then be shown to guarantee stack safety by exhibiting a fully abstract
%embedding function
%from the idealized machine to an ordinary machine equipped with this
%enforcement mechanism.  \bcp{Many OOPSLA readers/reviewers will not
%  understand that sentence.}
%%
%% While this work represents an important step in the right direction, we believe
%% there is still room to craft an operational characterization
%% of stack safety. \apt{Perhaps our rhetoric should be more micro-policy-specific here?}
%% In other words, we seek formal criteria that can be applied
%% to a mechanism for protecting stack-based low-level programs to judge
%% whether or not it actually ``guarantees stack safety.''
%% \rb{The remainder of this paragraph may be diving too quickly, would a reader be
%% confused by the references to these types of properties, yet to be introduced?}
%% We begin with a strong
%% local state encapsulation property that closely models the behavior of eager
%% micro-policies from \citet{DBLP:conf/sp/RoesslerD18}. This ``eager'' property
%% is too strong to characterize stack safety; reasonable enforcement mechanisms
%% exist that violate it. One such mechanism is a ``lazy'' tag policy, similar to
%% the one in \citet{DBLP:conf/sp/RoesslerD18}. We weaken local state encapsulation
%% to support this technique and argue that the new property, defined relative to an
%% arbitrary notion of observation. This property is still strong enough to prevent
%% the bad behaviors we associate with unsafe stack behavior, but weak enough that
%% other reasonable enforcement mechanisms can enforce it.

In the present work, we also aim to protect both stack data (like
LSE) and
control flow (like WBCF).  Our technical formulation of WBCF is similar to
that of Skorstengaard et al.{}, but our treatment of stack data protection
is rather different.  Instead of defining it implicitly,
as ``what a certain idealized machine does,'' we give an
\emph{explicit} definition using the technical framework of language-based
security\ifaftersubmission~\citep{??}\bcp{Maybe the Sabelfeld and Myers
  survey?}\fi, phrased as the conjunction of
\emph{stack integrity} and \emph{stack confidentiality}.
%
A key technical novelty in these definitions,
compared to standard formulations of confidentiality and integrity from the
security literature, is that they
are ``nested'': {\em each} caller is guaranteed protection from its
immediate callees (which, in turn, need protection from their immediate
callees, etc.). They are also stronger than LSE; even in the absence of the
control flow guarantees of {\em return to caller}, they guarantee callers'
data safety in the face of control-flow attacks (see \cref{sec:control}).
\rb{Reference to \cref{sec:running-example}, where this is discussed now?}
This reflects an arguably cleaner division between data protection and
control-flow protection than LSE / WBCF.

For the confidentiality and integrity properties, we consider both {\em
  stepwise} and {\em observational} variants.  The simpler {stepwise}
variants formalize the intuition that a caller's local data is {\em never}
read or modified during a call. The more extensional observational variants
allow a callee to read from and write to their caller's stack frame, as
long as nothing that they read affects their observable behavior and nothing
they write affects the observable behavior of the caller after they return.

@@

We are finally ready for our definition of stack safety. We begin
by describing how we model knowledge about the program structure,
then dive into the formal definitions of local state encapsulation\apt{we don't want
  to call it that here, do we?} and
well-bracketed control flow, which together constitute stack safety.

We will focus on a simple stack model, with a single stack
that doesn't share between caller and callee: arguments and results are passed
in registers. To accommodate different machine architectures and different calling
conventions, we parameterize over the structure of the stack, particularly
the means by
which a caller identifies the data that should be protected during a call. We term
this the {\em sealing convention}, \(\sealcon \subseteq \MACHS \times \COMPONENTS\), a
relation between machine states and the addresses that should be sealed.
If a state's \(\PCname\) is annotated as a call, the sealing convention tells us
which addresses must be protected until the return. Similarly, we abstract over what
it means to return with a {\em return convention}, \(\retcon \subseteq \MACHS \times \MACHS\),
which relates call states to the states that can be considered valid returns to that call.
The usual return convention is that \(\mach\) and \(\mach'\) agree on the stack pointer,
and \(\mach'\) has its program counter advanced by one instruction from \(\mach\).

Each system component is assigned to a {\em domain}, which may be {\em Outside}
of the stack entirely, {\em Unsealed} and therefore accessible, or {\em Sealed}
at a particular depth. Each caller registers a {\em target}
condition---a predicate on machine states that must hold for it to consider itself active.
A context is a pair of a ``domain map'' from components to domains, and a
``return target list,'' which is a stack of targets, one for each caller awaiting
return.  \ifaftersubmission\bcp{spacing is funny}\fi
%
\[\stackDom \in \STKDOMS ::= \outside + \unsealed + \sealed{\depth} \hspace*{0.5in} \target \subseteq \MACHS\]
\[(\domMap,\rts) = \context \in \CONTEXTS ::= (\COMPONENTS \rightarrow \STKDOMS)
  \times (\listT (\target)) \]
%
The initial context \(\context_0 = \domMap_0, []\), where \(\domMap_0\) maps
the stack to \(\unsealed\) and all other addresses and all registers to \(\outside\).

@@

Naturally, testing also revealed a number of errors in our
implementation of the enforcement mechanism (the original was written in C++
and targeted ARM machine code;
%\bcp{right?}\leo{yeah}
we re-implemented it in Coq targeting RISC-V).  These errors range
from trivial typos to ones that require an intriguingly complex setup
to appear.  The most interesting bug (included in the table as row
{\em HEADER\_NO\_INIT}) was that, on our first try, the blessed call
sequence %/policy combination\apt{??}
did not initialize all locations for the
newly allocated stack frame correctly, but left some of them as
$\tagNoDepth$. This allowed for a potential integrity violation, but
only if a rather complicated sequence of events occured.
The smallest counterexample requires calling a function {\tt f},
which writes into an uninitialized location $l$ (which is always fine\apt{is it??? in the strict policy?}, and then
calls a further function {\tt g} (which should have the effect of sealing $l$
for integrity purposes). Then {\tt g} attempts to write to $l$;
this is allowed because the enforcement mechanism still has
$l$ tagged as $\tagNoDepth$, but violates the integrity property on {\tt f}'s data.
\section{Stack Safety for Lazy Enforcement}
\label{sec:lazy}

@@

First, they eliminate the initialization pass. In this {\em Lazy Tagging}
system, the stack is initially tagged as ``clear.''
\rb{The previous section already presents ``unused'' as the default tag,
  though it doesn't say that it can be overwritten whenever\apt{but see my
query re the HEADER\_NO\_INIT bug}}
A function is permitted
to overwrite clear memory, and when it does so its depth tag overwrites
the clear tag, claiming the memory; the remainder of the policy functions
as {\em Depth Isolation}. On return, the callee must clear its frame, which
is still a heavy cost.

To eliminate this cost, \citet{DBLP:conf/sp/RoesslerD18} augment Lazy Tagging
with {\em Lazy Clearing}. Under this combination of optimizations,
the policy does not initialize stack frames on entry \emph{or} clear them on exit,
and it permits all writes to the stack, even when the $\PCname$ tag
does not match the memory tag; the \(\PCname\) tag is always propagated
to the written location. Reads from the stack do still require $\PCname$
tag and memory tag to match. Thus, even if a callee illicitly writes to
a private location in its caller, the caller will eventually detect this
if it ever tries to read from that location.
\rb{This is in line with the level of detail in the last section.}
%
These \emph{lazy policies} admit more efficient implementations, but
they deliberately allow temporary violations of stack integrity,
with the checks deferred until the point a violation truly becomes
harmful. The natural question then is: how do we characterize the
protections provided by these policies once stepwise stack safety is broken?
How does a harmful violation that must be caught later differ from a
harmless one?

@@

We first notice that, in principle\apt{??}, these properties appear to enforce
confidentiality outright. So we wish in particular to find an integrity
property. The intuition is that sealed memory can change, but it is
then tainted, and its original owner cannot actually read it. This
leads us to a hybrid of integrity and confidentiality.

%For simplicity, our definitions return to the simple stack without sharing,
%but all properties in this section can be extended to more complex settings
%very naturally.

\definition
For any \(\MPCT,\brk \in \MPCT^*\), let \(\head(\MPCT) = \mach,\pol,\context\) and
suppose that \(\last(\MPCT) = \mach',\pol',\context'\) and
\((\mach',\pol',\context') \hookrightarrow \MPCT'\). Let 
\(\components = \prot(\context) \cap \Delta(\mach,\mach')\), the
set of sealed components that changed during the call.
A system enjoys \textbf{\textit{deferred stepwise integrity}} if, for any \(\nach'\) and \(\NPCT'\)
such that \(\mach' \approx_{\components} \nach'\) and 
\((\nach',\pol',\context') \hookrightarrow \NPCT'\), \(\MPCT' \doteq \NPCT'\) holds.\apt{again, shouldn't
we quantify over all $M$?}

In short, if sealed components are modified during the call, the caller may not
access their values.\apt{This is pretty cryptic.}
%
For example, consider again the integrity violation in
Figure~\ref{fig:int1}. Under a deferred discipline, the write to {\tt
  x} in instruction 11 taints the variable but does not in itself
constitute a violation, which happens only when the tainted value is
read in instruction 4. However, if this instruction added the constant
that was originally stored in {\tt x} without reading the variable,
i.e., {\tt add \$42,\%ra}, deferred integrity would not be violated.
Instead of strictly forbidding such dangerous writes, we say,
``No reads after dangerous writes.''

@@

We argue that while deferred stepwise integrity meaningfully captures the intuition
behind lazy tagging, it stops short of the kind of extensional characterization
that would be viable as a weakest reasonable definition. Consider a hypothetical
taint-tracking policy that allows dangerous writes to be copied as long as the
taint remains hidden from outside observers. In order to characterize
such a policy, we would need a model of observation.
Similarly, for such a taint-based policy, lockstep confidentiality would be too
strong, as we might permit secrets to move about as long as they were not
leaked to the outside world.

@@

\definition
For any \(\MPCT,\brk \in \MPCT^*\),
let \(\head(\MPCT) = \mach,\pol,\context\).
%
Then, supposing that \(\last(\MPCT) = \mach',\pol',\context'\),
let \(\components\) be the set of protected components that changed during
the call, \(\components = \prot(\context) \cap \Delta(\mach, \mach')\),
and let \((\mach,\pol',\context') \hookrightarrow \MPCT'\).
A system enjoys {\em observational integrity} if for any \(\nach\) such that
\(\mach' \approx_{\components'} \nach\), and its induced trace
\((\nach,\pol',\context') \hookrightarrow \NPCT\), \(\obsof(\MPCT') \simeq \obsof(\NPCT)\).
\apt{Adjust quantification to match final version of deferred intergrity.}

@@

\definition
For any \(\MPCT,\brk \in \MPCT^*\),
let \(\head(\MPCT) = \mach,\pol,\context\) and
\(\components = \secret(\context)\).
%
A system enjoys {\em observational confidentiality} if for all
\(\nach\) such that \(\mach \approx_\components \nach\) and
\(\nach \hookrightarrow \NPCT | \brk(\MPCT)\),
\(\MPCT\) and \(\NPCT\), the following conditions all hold:

\begin{itemize}
\item They are observationally equivalent, \(\obsof(\MPCT) = \obsof(\NPCT)\),
  meaning that no secrets leak during the call
\item \(\brk ~ \last(\MPCT) \leftrightarrow \brk ~
\last(\NPCT)\)---if one call returns, so does the other
\item Assuming both calls return, let \(\last(\MPCT) = \mach',\pol',\context'\)
  and \(\pi_\mach(\last(\NPCT)) = \nach'\).
  Let \(\components' = \components \cap \bar{\Diamond}(\mach,\mach',\nach,\nach')\).
  And let \(\nach''\) be a state such that \(\mach' \approx_\components \nach''\).
  Then if \(\mach',\pol',\context' \hookrightarrow \MPCT'\) and
  \(\nach',\pol',\context' \hookrightarrow \NPCT'\), it must be
  that \(\obsof(\MPCT') \simeq \obsof(\NPCT')\)\apt{adjust quantification if needed}
\end{itemize}

@@

Observable properties allow us to defer enforcement until a property
violation would become visible, as in lazy policies. But by formalizing
the connection to the eager policy\apt{???}, we can identify when lazy policies
miss violations that can become visible. The {\em Lazy Tagging and Clearing}
micro-policy from \citet{DBLP:conf/sp/RoesslerD18} tags each stack slot with the call depth
at which it was written, and enforces that it must be read from the
same depth. However, this allows violations of observable stack safety!
Specifically, a callee might write into the stack and leave tags behind
for a future callee at the same depth to read. This violates confidentiality and,
if the writes were in the caller's frame, integrity.

@@

Heap safety, like stack safety, may be framed as a security property
in the form of
noninterference~\citep{DBLP:conf/post/AmorimHP18}\ifaftersubmission\bcp{Should
be spelled ``Azedevo et al.''  Got to figure out how to make bibtex do that
(or ask arthur how!).}\fi. Just
as~\citeauthor{DBLP:conf/post/AmorimHP18} give a rigorous
characterization of the meaning of (heap) memory safety, in this paper,
inspired by their work, we aim to do the same for the stack.
%
Their model describing safety of shared stack-allocated
objects\BCP{huh?? they were not talking about stacks} extends naturally to heap safety, complicated by the fact that
these objects are not deallocated by returns\bcp{??}, and therefore a
function's privilege may increase or decrease after its entry
point. For instance, if a callee allocates a heap object and returns
the pointer to its caller, that object's addresses become accessible
in contradiction to a contour computed at the caller's entry. So such
an extension must enforce trace confidentiality and integrity
properties separately on continuous segments of each call, with
contours computed at each crossing between caller and callee.  \BCP{Didn't
  understand very much of that paragraph.}

>>>>

%\bcp{Does use of unitialized variables lead to stack safety attacks?\apt{sure, via buffer overflow}}
%with the notion usually being
%defined negatively: through examples of stack-based exploits
%\rb{usually though not always; as noted below there are other
%  approaches}\bcp{Yes, we have to be very careful about this}.
%\leo{Examples at a high level}: such as -- does not overwrite return addresses, or ...
The most abstract characterization of stack safety that we are aware of is
recent work by \citet{SkorstengaardSTK}, which defines stack safety as the conjunction of two
properties:
{\em local state encapsulation (LSE)} and {\em well-bracketed control flow (WBCF)}.
Informally, LSE says that the contents of
a caller's stack frame are not read or written while its callees (and their callees, etc.) are
executing, and WBCF says that callees always
return to the instruction following their call (if they return at all).
Skorstengaard et al.{} formalize these properties by defining an idealized
machine in which they hold {\em by construction};
they then show that an enforcement
mechanism (in their case, hardware capabilities) guarantees these properties, by
proving that an ordinary machine with enforcement simulates the idealized
machine.
%an enforcement mechanism (in their case, hardware
%capabilities)
%can then be shown to guarantee stack safety by exhibiting a fully abstract
%embedding function
%from the idealized machine to an ordinary machine equipped with this
%enforcement mechanism.  \bcp{Many OOPSLA readers/reviewers will not
%  understand that sentence.}
%%
%% While this work represents an important step in the right direction, we believe
%% there is still room to craft an operational characterization
%% of stack safety. \apt{Perhaps our rhetoric should be more micro-policy-specific here?}
%% In other words, we seek formal criteria that can be applied
%% to a mechanism for protecting stack-based low-level programs to judge
%% whether or not it actually ``guarantees stack safety.''
%% \rb{The remainder of this paragraph may be diving too quickly, would a reader be
%% confused by the references to these types of properties, yet to be introduced?}
%% We begin with a strong
%% local state encapsulation property that closely models the behavior of eager
%% micro-policies from \citet{DBLP:conf/sp/RoesslerD18}. This ``eager'' property
%% is too strong to characterize stack safety; reasonable enforcement mechanisms
%% exist that violate it. One such mechanism is a ``lazy'' tag policy, similar to
%% the one in \citet{DBLP:conf/sp/RoesslerD18}. We weaken local state encapsulation
%% to support this technique and argue that the new property, defined relative to an
%% arbitrary notion of observation. This property is still strong enough to prevent
%% the bad behaviors we associate with unsafe stack behavior, but weak enough that
%% other reasonable enforcement mechanisms can enforce it.

In the present work, we also describe the protection of both stack data (like LSE) and
control flow (like WBCF).  Our technical formulation of WBCF is similar to
theirs, but our treatment of stack data protection is rather different,
as is our approach to the definition itself. Instead of defining stack safety implicitly,
as ``what a certain idealized machine does,'' we give an
\emph{explicit} definition using the technical framework of language-based
security\ifaftersubmission~\citep{??}\bcp{Maybe the Sabelfeld and Myers
  survey?}\fi, phrased as the conjunction of
\emph{stack integrity} and \emph{stack confidentiality}.
%
A key technical novelty in these definitions,
compared to standard formulations of confidentiality and integrity from the
security literature, is that they
are ``nested'': {\em each} caller is guaranteed protection from its
immediate callees (which, in turn, need protection from their immediate
callees, etc.). They are also stronger than LSE; even in the absence of the
control flow guarantees of WBCF, they guarantee callers'
data safety in the face of control-flow attacks (see \cref{sec:controlflow}).
\rb{Reference to \cref{sec:running-example}, where this is discussed now?}
This reflects an arguably cleaner division between data protection and
control-flow protection than LSE / WBCF.

For confidentiality and integrity , we consider both {\em stepwise}
and {\em observational} variants.  The simpler stepwise
variants formalize the intuition that a caller's local data is {\em never}
read or modified during a call. The observational variants
allow a callee to read from and write to their caller's stack frame, as
long as nothing that they read affects their observable behavior and nothing
they write affects the observable behavior of the caller after they return.
The observational properties are meant to capture the same sense of security,
extensionally enough that any reasonable protection mechanism ought to enforce
them, even if it does not prevent every single dangerous read or write.

@@

We are finally ready for our definition of stack safety. We begin
by describing how we model knowledge about the program structure,
then dive into the formal definitions of integrity, confidentiality, and
well-bracketed control flow, which together constitute stack safety.

We will focus on a simple stack model, with a single stack
that doesn't share between caller and callee: arguments and results are passed
in registers. To accommodate different machine architectures and different calling
conventions, we parameterize over the structure of the stack, particularly
the means by
which a caller identifies the data that should be protected during a call. We term
this the {\em sealing convention},
\(\sealcon ~ (\mach \in \MACHS) ~ (\component \in \COMPONENTS)\), a
relation between machine states and the addresses that should be sealed.
If a state's \(\PCname\) is annotated as a call, the sealing convention tells us
which addresses must be protected until the return. Similarly, we abstract over what
it means to return with a {\em return convention},
\(\retcon ~ (\mach \in \MACHS) ~ (\mach' \in \MACHS)\),
which relates call states to the states that can be considered valid returns to that call.
The usual return convention is that \(\mach\) and \(\mach'\) agree on the stack pointer,
and \(\mach'\) has its program counter advanced by one instruction from \(\mach\).

Each system component is assigned to a {\em domain}, which may be {\em Outside}
of the stack entirely, {\em Unsealed} and therefore accessible, or {\em Sealed}
at a particular depth. Each caller registers a {\em target}
condition, \(\target ~ \mach\)---a predicate on machine states that means it
has been returned to.
A context is a pair of a ``domain map'' from components to domains, and a
``return target list,'' which is a stack of targets, one for each caller awaiting
return.
%
\[\stackDom \in \STKDOMS ::= \outside + \unsealed + \sealed{\depth} \hspace*{0.5in} \target \subseteq \MACHS\]
\[(\domMap,\rts) = \context \in \CONTEXTS ::= (\COMPONENTS \rightarrow \STKDOMS)
  \times (\listT (\target)) \]
%
The initial context \(\context_0 = \domMap_0, []\), where \(\domMap_0\) maps
the stack to \(\unsealed\) and all other addresses and all registers to \(\outside\).

@@

Naturally, testing also revealed a number of errors in our
implementation of the enforcement mechanism (the original was written in C++
and targeted ARM machine code;
%\bcp{right?}\leo{yeah}
we re-implemented it in Coq targeting RISC-V).  These errors range
from trivial typos to ones that require an intriguingly complex setup
to appear.  The most interesting bug (included in the table as row
{\em HEADER\_NO\_INIT}) was that, on our first try, the blessed call
sequence %/policy combination\apt{??}
did not initialize all locations for the
newly allocated stack frame correctly, but left some of them as
$\tagNoDepth$. This allowed for a potential integrity violation, but
only if a rather complicated sequence of events occured.
The smallest counterexample requires calling a function {\tt f},
which writes into an uninitialized location $l$ (which is always fine\apt{is it??? in the strict policy?}, and then
calls a further function {\tt g} (which should have the effect of sealing $l$
for integrity purposes). Then {\tt g} attempts to write to $l$;
this is allowed because the enforcement mechanism still has
$l$ tagged as $\tagNoDepth$, but violates the integrity property on {\tt f}'s data.
\sna{I believe what went wrong was that we were off-by-one in {\tt main}'s initialization,
  and the write from {\tt f} was already a violation.}

\section{Stack Safety for Lazy Enforcement}
\label{sec:lazy}

@@

First, they eliminate the initialization pass. In the {\em Lazy Tagging}
system, the stack is initially tagged as \(\tagNoDepth\),
\rb{The previous section already presents ``unused'' as the default tag,
  though it doesn't say that it can be overwritten whenever\apt{but see my
    query re the HEADER\_NO\_INIT bug}}\sna{Correct there, but now I'm talking
  about laziness.}
A function is permitted
to overwrite clear memory, and when it does so its depth tag overwrites
the clear tag, claiming the memory; the remainder of the policy functions
as {\em Depth Isolation}. On return, the callee must clear its frame, restoring
it to \(\tagNoDepth\), which is still a heavy cost.

To eliminate this cost, \citet{DBLP:conf/sp/RoesslerD18} augment Lazy Tagging
with {\em Lazy Clearing}. Under this combination of optimizations,
the policy does not initialize stack frames on entry \emph{or} clear them on exit,
and it permits all writes to the stack, even when the $\PCname$ tag
does not match the memory tag; the \(\PCname\) tag is always propagated
to the written location. Reads from the stack do still require $\PCname$
tag and memory tag to match. Thus, even if a callee illicitly writes to
a private location in its caller, the caller will eventually detect this
if it ever tries to read from that location.
%
These \emph{lazy policies} admit more efficient implementations, but
they deliberately allow temporary violations of stack integrity,
with the checks deferred until the point a violation truly becomes
harmful. The natural question then is: how do we characterize the
protections provided by these policies once stepwise stack safety is broken?
How does a harmful violation that must be caught later differ from a
harmless one?

@@

We first notice that these properties prevent invalid reads, as well as reads
that follow invalid writes, but the writes themselves are permitted.
So we wish in particular to find an integrity property that characterizes them.
The intuition is that sealed memory can change, but it is
then tainted, and its original owner cannot actually read it. This
leads us to a hybrid of integrity and confidentiality.

\definition
A system enjoys \textbf{\textit{deferred stepwise integrity}} if,
for any \(\MPCT,\brk \in \MPCT^*\) where \(\head(\MPCT) = \mach,\pol,\context\) and
\(\last(\MPCT) = \mach',\pol',\context'\), the following holds. Let
\((\mach',\pol',\context') \hookrightarrow \MPCT'\) and
\(\components = \prot(\context) \cap \Delta(\mach,\mach')\), the
set of sealed components that changed during the call.
For any \(\nach'\) and \(\NPCT'\) such that \(\mach' \approx_{\components} \nach'\) and 
\((\nach',\pol',\context') \hookrightarrow \NPCT'\), \(\MPCT' \doteq \NPCT'\) holds.

For example, consider again the integrity violation in
Figure~\ref{fig:int1}. Under a deferred discipline, the write to {\tt
  x} in instruction 11 taints the variable but does not in itself
constitute a violation, which happens only when the tainted value is
read in instruction 4. However, if this instruction added the constant
that was originally stored in {\tt x} without reading the variable,
i.e., {\tt add \$42,\%ra}, deferred integrity would not be violated.
Instead of strictly forbidding such dangerous writes, we say,
``No reads after dangerous writes.''

@@

While deferred integrity meaningfully captures the intuition
behind lazy tagging, we seek properties that can be used to evaluate any reasonable
enforcement mechanism. Consider a hypothetical
taint-tracking policy that allows high-security to be copied as long as it is
tagged as such and not exfiltrated, or that allows dangerous writes to be
copied as long as they don't affect critical operations. This calls for
an abstract mechanism to characterize behaviors that are externally relevant.

@@

\definition
{\em Observational integrity} means that for any \(\MPCT,\brk \in \MPCT^*\)
where \(\head(\MPCT) = \mach,\pol,\context\) and \(\last(\MPCT) = \mach',\pol',\context'\),
the following holds. Let \(\components\) be the set of protected components that changed during
the call, \(\components = \prot(\context) \cap \Delta(\mach, \mach')\),
and let \((\mach,\pol',\context') \hookrightarrow \MPCT'\).
Then for any \(\nach\) such that
\(\mach' \approx_{\components'} \nach\), and its induced trace
\((\nach,\pol',\context') \hookrightarrow \NPCT\), \(\obsof(\MPCT') \simeq \obsof(\NPCT)\).

@@

\definition
A system enjoys {\em observational confidentiality} if for all
\(\MPCT,\brk \in \MPCT^*\), with \(\head(\MPCT) = \mach,\pol,\context\) and
\(\components = \secret(\context)\), the following holds.
%
For any \(\nach\) such that \(\mach \approx_\components \nach\) and
\(\nach \hookrightarrow \NPCT | \brk(\MPCT)\),
\(\MPCT\) and \(\NPCT\) respect three conditions:

\begin{itemize}
\item They are observationally equivalent, \(\obsof(\MPCT) = \obsof(\NPCT)\),
  meaning that no secrets leak during the call
\item \(\brk ~ \last(\MPCT) \leftrightarrow \brk ~
\last(\NPCT)\)---if one call returns, so does the other
\item Assuming both calls return, let \(\last(\MPCT) = \mach',\pol',\context'\)
  and \(\pi_\mach(\last(\NPCT)) = \nach'\).
  Let \(\components' = \components \cap \bar{\Diamond}(\mach,\mach',\nach,\nach')\).
  And let \(\nach''\) be a state such that \(\mach' \approx_\components \nach''\).
  Then if \(\mach',\pol',\context' \hookrightarrow \MPCT'\) and
  \(\nach',\pol',\context' \hookrightarrow \NPCT'\), it must be
  that \(\obsof(\MPCT') \simeq \obsof(\NPCT')\)
\end{itemize}

@@

Observable properties allow us to defer enforcement until a property
violation would become visible. Now we can identify when lazy policies
miss violations that can become visible. The {\em Lazy Tagging and Clearing}
micro-policy from \citet{DBLP:conf/sp/RoesslerD18} tags each stack slot with the call depth
at which it was written, and enforces that it must be read from the
same depth. However, this allows violations of observable stack safety!
Specifically, a callee might write into the stack and leave tags behind
for a future callee at the same depth to read. This violates confidentiality and,
if the writes were in the caller's frame, integrity.

@@

Heap safety, like stack safety, may be framed as a security property
in the form of
noninterference~\citep{DBLP:conf/post/AmorimHP18}\ifaftersubmission\bcp{Should
be spelled ``Azedevo et al.''  Got to figure out how to make bibtex do that
(or ask arthur how!).}\fi. Just
as~\citeauthor{DBLP:conf/post/AmorimHP18} give a rigorous
characterization of the meaning of (heap) memory safety, in this paper,
inspired by their work, we aim to do the same for the stack. They are a
significant inspiration for this work.
